---
title: "Efficacy of Transdiagnostic Cognitive-Behavioral Therapy for Assertiveness: A Randomized Controlled Trial"
shorttitle: "Efficacy of CBT for Assertiveness"
author:
- name: Tobias Hagberg
  affiliation: SU
- name: Patrik Manhem
  affiliation: SU
- name: Martin Oscarsson
  affiliation: SU
- name: Fiona Michel
  affiliation: CCI
- name: Gerhard Andersson
  affiliation: LU,KI
- name: Per Carlbring
  email: per.carlbring@psychology.su.se
  correspondingauthor: yes
  affiliation: SU
date: ''
output:
  #bookdown::word_document2:
  papaja::apa6_docx:
    number_sections: false
    tables:
      style: Table
      layout: autofit
      width: 1.0
      caption:
       style: Table Caption
       pre: "Table "
       sep: ". "
    plots:
      style: Normal
      layout: autofit
      caption:
       style: Image Caption
       pre: "Figure "
       sep: ". "
abstract: "`r readLines('abstract.Rmd', encoding = 'UTF-8')`\n"
keywords:
- assertiveness
- assertive behavior
- anxiety
- depression
- stress
- avoidance
runtitle: Transdiagnostic CBT for Assertiveness
bibliography:
- packages.bib
- KSH.bib
- DBT.bib
csl: "./elsevier-harvard.csl"
address:
- code: SU
  address: "Department of Psychology, Stockholm University, SE-106 91 Stockholm, Sweden"
- code: KI
  address: "Department of Clinical Neuroscience, Karolinska Institutet, SE-171 77
    Stockholm, Sweden"
- code: LU
  address: "Department of Behavioural Sciences and Learning, Department of Biomedical
    and Clinical Sciences, Linköping University, SE-581 83 Linköping, Sweden"
- code: CCI
  address: Centre for Clinical Interventions, 223 James St, Northbridge WA 6003, Australia
---

## Highlights

  - Stand-alone iCBT assertiveness training is effective at increasing levels of adaptive assertiveness.
  - Stand-alone iCBT assertiveness training is effective at alleviating symptoms of social anxiety to the point of clinical recovery for a significant number of participants.
  - The effects were maintained at the one year follow-up.
  - The assertiveness construct can also be useful as a target in a self-help guide, without therapist support.
  - The assertiveness construct is worthy of revived clinical research interest.

```{r setup, include=F}

# https://hughjonesd.github.io/huxtable/reference/label.html
# https://ardata-fr.github.io/officeverse/index.html
# https://gt.rstudio.com/

packages <-
  c(
    "pwr",
    "readxl",
    "writexl",
    "lmerTest",
    "dunn.test",
    "ggplot2",
    "tidyr",
    "jtools",
    "modelr",
    "ggeffects",
    "emmeans",
    "patchwork",
    "performance",
    "sjPlot",
    "clinicalsignificance",
    "here",
    "linebreak",
    "MOTE",
    "rticles",
    "devtools"
  )
if (!require("install.load")) install.packages("install.load")
library(install.load)
install_load(packages)

# https://github.com/haozhu233/kableExtra/issues/750
devtools::install_github("kupietz/kableExtra")
library(kableExtra)

knitr::opts_chunk$set(
  warning = F,
  message = F,
  echo = F,
  collapse = T,
  fig.retina = 4,
  dpi = 196,
  out.width = "100%",
  fig.width = 7.5,
  fig.asp = 0.618
)

```

```{r body, child="body-w.Rmd", eval=TRUE}
```